---
title: "JSC370 Final Project Interactive Visualizations"
# author: "Tsai-Ni Lin"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---


```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(cache=TRUE, echo=FALSE, warning = FALSE, message = FALSE)
```


```{r,echo=FALSE}
library(data.table)
library(tidyverse)
library(plotly)
library(knitr)
library(widgetframe)
library(dplyr)
library(ggpubr)
library(ggplot2)
library(magrittr)
library(kableExtra)
library(scales)
library(xgboost)
library(gbm)
library(leaflet)
library(caret) 
library(gamair)
library(mgcv)
library(randomForest)
library(leaflegend)
library(ggpubr)
```

```{r}
data <- readRDS('../data/midterm_data.rds')
```

```{r}
#load data
leaflet_marker <- readRDS('../data/leaflet_marker.rds')
sunshine_by_city <- readRDS('../data/sunshine_by_city.rds')
sunshine_data <- readRDS('../data/sunshine_data.rds')
```


```{r}
#records after 2010
country_records2 <- data %>%  filter(Year >= 2010) %>% group_by(Country) %>% mutate(record_count = n())

#maximum amount of time country was reported
max_count <- max(country_records2$record_count)
  
#filter out countries that don't appear for "max count of times"  after 2010
freq_countries <- country_records2 %>% filter(record_count >= max_count) 


#yearly mean happiness in selected countries
main_plt_data <- freq_countries %>% group_by(Year) %>%summarise(mean_happiness = mean(Happiness), level = "Overall", n=n())


#mean of gdp, sunshine, and freedom
mean_gdp = mean(freq_countries$Log_GDP)
mean_sunshine = mean(freq_countries$Avg_sunshine)
mean_freedom = mean(freq_countries$Freedom)

#create income category 
levels_countries_income <- freq_countries %>% mutate(level = case_when(Log_GDP >= mean_gdp 
                                                   ~ "Income Above Average", TRUE ~"Income Below Average" )) 

#calculate the mean happiness in each year by income category
plt_income <- levels_countries_income %>% group_by(Year, level) %>% summarise(mean_happiness = mean(Happiness), n=n())



#create sunshine category 
levels_countries_sunshine <- freq_countries %>% mutate(level = case_when(Avg_sunshine >= mean_sunshine 
                                                   ~ "Sunshine Above Average", TRUE ~"Sunshine Below Average" )) 



#calculate the mean happiness score in each year by sunshine category
plt_sunshine <- levels_countries_sunshine %>% group_by(Year, level) %>% summarise(mean_happiness = mean(Happiness), n=n())



# create freedom category
levels_countries_freedom <- freq_countries %>% mutate(level = case_when(Freedom >= mean_freedom 
                                                   ~ "Freedom Above Average", TRUE ~"Freedom Below Average" )) 

#calculate the mean happiness score in each year by freedom category
plt_freedom <- levels_countries_freedom %>% group_by(Year, level) %>% summarise(mean_happiness = mean(Happiness), n=n())


#combine the data frames
time_plot <- bind_rows(plt_income, plt_sunshine, plt_freedom, main_plt_data)
```


```{r}
#plot time series plot
 p2 <- ggplot(data=time_plot, aes(x=Year,y = mean_happiness, color = level)) +
geom_line() + 
  geom_point(data=time_plot, aes(x=Year,y = mean_happiness, text= paste("Year: ", Year, "<br>", 
                                    "Mean Happiness Score: ", round(mean_happiness,3), "<br>",
                                    "Level:", level, "<br>", sep = " "
                                    ))) + guides(fill=guide_legend(title="Country Category")) + 
   labs(y = "Mean Happiness Score", x = "Year", title = "Mean Happiness Score Overtime for Countries recorded every year bewteen 2010-2020") + theme(plot.title = element_text(size=10))

#adjust x-axis scales
 p2 <- p2+ 
scale_x_continuous(breaks= c(2010, 2015, 2020))
```

```{r}
sum <- data%>% group_by(Country) %>% mutate(avg_gdp = mean(Log_GDP), avg_freedom = mean(Freedom)) 

mean_freedom <- mean(data$Freedom)
sum <- sum %>% mutate(freedom_level = case_when(Freedom >= mean_freedom 
                                                   ~ "Freedom Above Average", TRUE ~"Freedom Below Average"))
```



```{r}
#linear - gdp
plt_smooth1 <- ggplot(
  sum, 
  aes(x = Log_GDP, y= Happiness,text= paste("Country: ", Country, "<br>", 
                                    "Log GDP: ", round(Log_GDP,3), "<br>",
                                    "Happiness Score: ", round(Happiness,3)))) + 
  theme_minimal() + 
  geom_smooth(inherit.aes = FALSE, aes(x = Log_GDP, y= Happiness)) + 
  geom_point(aes(colour = Country),alpha = 4/10) + 
labs(x = "Log GDP", y = "Happiness Score", title = "Happiness Score vs. Log GDP by Country")
```


```{r}
#linear - freedom
plt_smooth2 <- ggplot(
  sum, 
  aes(x = Freedom, y= Happiness,text= paste("Country: ", Country, "<br>", 
                                    "Freedom Score: ", round(Freedom,3), "<br>",
                                    "Happiness Score: ", round(Happiness,3)))) + 
  theme_minimal() + 
  geom_smooth(inherit.aes = FALSE, aes(x = Freedom, y= Happiness)) + 
  geom_point(aes(colour = Country),alpha = 4/10) + 
labs(x = "Log GDP", y = "Happiness Score", title = " Happiness Score vs. Freedom Score by Country")
```


```{r}
#linear - sunshine
plt_smooth3 <- ggplot(
  sum, 
  aes(x = Avg_sunshine, y= Happiness,text= paste("Country: ", Country, "<br>", 
                                    "Sunshine Hours: ", round(Avg_sunshine,3), "<br>",
                                    "Happiness Score: ", round(Happiness,3)))) + 
  theme_minimal() + 
  geom_smooth(inherit.aes = FALSE, aes(x = Avg_sunshine, y= Happiness)) + 
  geom_point(aes(colour = Country),alpha = 4/10) + 
labs(x = "Sunshine Hours", y = "Happiness Score", title = " Happiness Score vs. Sunshine Hours by Country")
```

```{r}

#distribution plots
plt1 <- ggplot(data, aes(x=Log_GDP)) + geom_histogram() + labs(xlab = 'Log GDP per Capita', title = 'Distribution of Log GDP per Capita')

plt2 <- ggplot(data, aes(x=Avg_sunshine)) + geom_histogram() + labs(xlab = 'Log GDP per Capita', title = 'Distribution of Yearly Sunshine Hours (hrs)')

plt3 <- ggplot(data, aes(x=Freedom)) + geom_histogram() + labs(xlab = 'Log GDP per Capita', title = 'Distribution of Freedom Score')

plt4 <- ggplot(data, aes(x=Happiness)) + geom_histogram() + labs(xlab = 'Log GDP per Capita', title = 'Distribution of Happiness Score')

```


```{r}
#create income level
data3 <- data %>% mutate(income_level = case_when((Log_GDP <= 
                                                   quantile(Log_GDP, probs =0.25, na.rm = TRUE)) ~ "Low",
                                                 (quantile(Log_GDP, probs =0.25, na.rm = TRUE) < Log_GDP & Log_GDP<= 
                                                   quantile(Log_GDP, probs =0.75, na.rm = TRUE)) ~ "Medium",
                                                TRUE ~ "High"))


#create sunshine level
data3 <- data3 %>% mutate(sunshine_level = case_when((Avg_sunshine < 
                                                   quantile(data$Avg_sunshine, probs =0.5, na.rm = TRUE)) ~ "Low Sunshine Level",TRUE ~ "High Sunshine Level"))


#create freedom level
data3 <- data3 %>% mutate(freedom_level = case_when((Freedom < 
                                                   quantile(data$Freedom, probs =0.5, na.rm = TRUE)) ~ "Low Freedom Level",TRUE ~ "High Freedom Level"))


#reorder x-axis
data3$income_level <- factor(data3$income_level , levels=c("High", "Medium", "Low"))
data3$sunshine_level <- factor(data3$sunshine_level , levels=c("High Sunshine Level", "Low Sunshine Level"))

data3$freedom_level <- factor(data3$freedom_level , levels=c("High Freedom Level", "Low Freedom Level"))



```



```{r}
#income level plt                                              
l1 <- ggplot(data = data3, mapping = aes(x = income_level, y = Happiness)) + 
geom_boxplot(fill="pink")+ labs(x = "Income Level", y= "Happiness Score") 

#sunshine_level plt                                              
l2 <- ggplot(data = data3, mapping = aes(x = sunshine_level, y = Happiness)) + 
geom_boxplot(fill="mediumpurple1") + labs(x = "Sunshine Level", y= "Happiness Score") + scale_x_discrete(labels=c("High Sunshine Level" = "High", "Low Sunshine Level" = "Low"))


#freedom level plt                                              
l3 <- ggplot(data = data3, mapping = aes(x = freedom_level, y = Happiness)) + 
geom_boxplot(fill="lavenderblush1") + labs(x = "Freedom Level", y= "Happiness Score") + scale_x_discrete(labels=c("High Freedom Level" = "High", "Low Freedom Level" = "Low"))

```



```{r, fig.show='hide'}
#create boxplot of happiness score that groups by sunshine, freedom, and income level
level_plt <- ggplot(data = data3, mapping = aes(fill = income_level, y = Happiness)) + 
geom_boxplot() + facet_wrap(~sunshine_level + freedom_level) + labs(x = "Income Level", y= "Happiness Score", title = "Distribution of Happiness Score based on Sunshine Level, Freedom Level, and Income Level (Combined)") + theme(plot.title = element_text(size=10)) + scale_fill_manual(values=c("purple",
                               "hotpink1", "yellow"
                               )) + guides(fill=guide_legend(title="Income Level"))


level_plt
```




## Figure 1: 


This visualization shows the sunshine hours of each city within the sunshine data set that is used in the project. The markers are grouped in high, medium, and low sunshine level, and these are the representational cities for each country. In other words, the country uses that cities' sunshine hours for analysis. 

```{r}

#palette
rh_pal <- colorNumeric(
  # c('lightgrey','lightblue','blue'),
  viridisLite::viridis(5, direction = -1),
  domain = sunshine_data$Year
  )

# icon for sunshine level marker 
icons <- awesomeIcons(
  icon = 'ios-close',
  iconColor = 'white',
  library = 'ion',
  markerColor = leaflet_marker$sunshine_level
)

# icon set for creating legend in leaflet 
iconSet <- awesomeIconList(
  `Low` = makeAwesomeIcon(icon = 'ios-close',
  iconColor = 'white',
  library = 'ion',
  markerColor = "pink",
  ),
  Medium = makeAwesomeIcon(icon = 'ios-close',
  iconColor = 'white',
  library = 'ion',
  markerColor = "red",
  ),
  High = makeAwesomeIcon(icon = 'ios-close',
  iconColor = 'white',
  library = 'ion',
  markerColor = "blue",
  ))

```

```{r}
# leaflet map for sunshine hours in countries
leaflet() %>%
  addProviderTiles('OpenStreetMap') |>
  addCircles(lat = ~lat, lng = ~lng, color = ~rh_pal(Year),
             label = ~paste0(City, ': ', round(Year,2), ' hrs\n'),
             opacity = 1, fillOpacity = 0.8, radius = 600, data = sunshine_by_city) |>
  addAwesomeMarkers(lat = ~lat, lng = ~lng, icon=icons, data = leaflet_marker, label = ~paste0(leaflet_marker$City, ': ', round(leaflet_marker$Avg_sunshine,2), ' hrs\n'))|>
  addLegend('bottomleft', pal = rh_pal, values = sunshine_by_city$Year,
            title = "Yearly Sunshine Hours", opacity=1)|>

  addLegendAwesomeIcon(iconSet = iconSet, title = 'Sunshine Level',
                       position = 'topleft',
                       group = 'Awesome Icons')
```




## Figure 2: 

This visualization shows the mean happiness score overtime in countries that participated in the Happiness report every year between 2010 to 2020. Trends are separated by income level, freedom level, and sunshine level. "Overall" is the trend for all these countries without separating them into distinct income, freedom, and sunshine levels. 



```{r echo=FALSE}
ggplotly(p2, tooltip = "text")
```





## Figure 3: {.tabset}


These figures demonstrate the distribution of happiness score and GDP levels, freedom levels, and sunshine hours levels. Income Levels are separated in 3 groups: low is below Q1, medium is Q1 - Q3, and high is above Q3, while freedom and sunshine levels are in 2 groups. 

**Note**: Please hit "autoscale" to better visualize the combined interactive boxplot.

### Income Level

```{r}
ggplotly(l1)
```

### Sunshine Level


```{r}
ggplotly(l2)
```

### Freedom Level

```{r}
ggplotly(l3)
```


### Combinations of Levels

```{r}
ggplotly(level_plt, dynamicTicks = FALSE, height = 1000, width = 800) %>%
    layout(xaxis = list(autorange = TRUE), yaxis = list(autorange = TRUE), title = list(y = .95, xref = "plot"),
    margin = list(r= 40, t = 150)
    )
```



{-}

## Figure 4: {.tabset} 


These figures demonstrate the relationship between happiness score and GDP, freedom score, and sunshine hours separated by country. 


### Log GDP

```{r}
ggplotly(plt_smooth1, tooltip = "text")
```
### Freedom Score

```{r}
ggplotly(plt_smooth2, tooltip = "text")
```

### Sunshine Hours

```{r}
ggplotly(plt_smooth3, tooltip = "text")
```


{-}

